Musical Instrument Classification Using a Hybrid Neural Network

### 2021-08-13 JMMPaper/PaperDrafts/main.tex
Major Updates to Section 2 and 3

By connecting multiple layers in succession, we form a chain-like structure of layers that describes the flow of information in the neural network. This allows us to model a neural network as a directed computational graph where each node is a layer that performs an operation  \cite{Geron,Goodfellow}. In a feed-forward network with $L$ layers, inputs $x^{(0)}$ and outputs $x^{(L-1)} = y^*$, we can represent a layer-chain such as:
\begin{equation}
    \label{eqn:LayerChain}
    x^{(0)} \rightarrow f^{(0)} \rightarrow f^{(1)} \rightarrow ... \rightarrow f^{(L-2)} \rightarrow f^{(L-1)} \rightarrow y^*
\end{equation}
The structure of a neural network's layer-chain defines it's \textit{architecture}. The choice of architecture indicates the \textit{hypothesis-space} of the model a determines the set of all possible solutions that a trained network can attain \cite{Goodfellow,James}. For deep neural networks this function chain can be dozens of layers long, and contain upwards of millions of parameters. Increasingly complicated networks allow for a high dimensional solution spaces, which potentially enables the model to describe more complicated systems.

The hybrid neural network used in this instrument classification task differs from the more standard architecture in Eq. (\ref{eqn:LayerChain}) in that the function chain consists of two input layers, each of which feed into their own respective layer chains, noted by $f_a$ and $f_b$. Each chain transforms activations through a non-overlapping graph, and are then merged at a subsequent hidden layer to produce graph $f_c$. The aggregation of both graphs allows us to generate a single prediction, $y^*$ based on the behavior of both inputs, $x^{(0)}_a$ and $x^{(0)}_b$. We detail the structure an function of each layer chain in section (\ref{subsec:CNN}) and (\ref{subsec:MLP}). We represent the layer-chain structure for a generalized double-input HNN, with inputs $x^{(0)}_a$ and $x^{(0)}_b$ and outputs $y^*$ as :

@@

This structure allows us to consider and design each branch $a$ and $b$ as if they were separate neural networks. The outputs from $f^{(\alpha)}_a$ and $f^{(\beta)}_b$ can be chosen to represents the predictions of the branches respectively, which are concatenated at layer $f^{(\gamma)}_c$. This marks the entry point of a new chain, which ends at the final output of the full model. Although we may consider the each layer-chain to be it's own entity, it is more appropriate to consider the chains to be subsets of the same computational graph. This means that Eq. (\ref{eqn:LayerGraph}) is considered to be one hybrid neural network model in the training and validation process.

@@

The inputs to a neural network or any machine learning algorithm are called \textit{features} or \textit{predictors} \cite{James}. These are compact, low-dimensional representations of a data sample that reflect its important characteristics \cite{Liu}. Features should be chosen as to have low variance within each class and high variance between classes. For digital audio classification of musical instruments, features of the same musical instruments should exhibit very similar properties, while features from difference musical instruments should exhibit non-similar properties. In any learning algorithm, the role of feature extraction is to translate the raw information from a digital audio file into descriptors that maximize the classification performance \cite{Virtanen}. Although the neural network will classify the musical instrument within the digital audio sample, it will never interact with the raw waveform directly, instead will rely solely on these features. Suppose we implement a classification model that uses $p$ unqiue predictors. These features are usually combined into a single vector-like object for each sample, and then presented as input to the neural network \cite{Geron}. We detail the classification features for this neural network in section (\ref{subsec:FeaturesMLP}) and (\ref{subsec:FeaturesCNN}).

The outputs to a classification neural network encodes the prediction that the model has made given a particular set of inputs from a sample. For a classifier with $k$ classes, there are $k$ output neurons transformed such that the sum (or $L-1$ norm) of the output activations is identically $1$ \cite{Goodfellow,James}. This means the $k$ neurons can be treated as a probability density for each of the classes, where the neuron with the highest activation value is the prediction of the sample. For classification of musical instruments, each class represents a musical instrument type. We use $37$ classes of instruments encapsulating woodwinds, brass, strings, mallet percussion, and synthesizers. Each audio file contains exactly one musical instrument sustaining one note.

@@

A neural network is trained by providing a batch of input samples $X$, with a corresponding set of expected predictions $Y$. Consider a sample $x \in X$ that is expected to produce corresponding label $y \in Y$. When passed into the network, $x$ produces some prediction label denoted as $y^*$. For an untrained network, we anticipate that $y^*$ and $y$ may differ greatly - meaning that the prediction is very different that the expected output, and classification performance is likely to be very poor. Since we know each sample in the data set to contain one, and only one musical instrument, we \textit{one-hot-encode} the expected label $y$. For any sample $x^{(i)}$ that belongs to class $j$, we have a corresponding target vector $y^{(i)}$ given by:
\begin{equation}
    \label{eqn:OneHotEnc}
    y^{(i)} = \big[ y_{0} , y_{1}, ... , y_{j}, ... , y_{k-1} \big] = \big[ 0, 0, ... , 1, ... , 0 \big]
\end{equation}
The neural network will output an equivalently sized vector, $y^{*(i)}$, which all elements sum to $1$.

To quantify the difference between $y$ and $y^*$, we introduce a \textit{cost function}, $J(y,y^*)$ which is also called an \textit{objective function} or a \textit{loss function} \cite{James}. The exact cost function used can differ based on the nature of the machine learning model or task, but typically multiple - category classification tasks use a \textit{categorical cross-entropy} (CXE) function \cite{Geron,Goodfellow}. The CXE cost for a single sample is defined as \cite{Goodfellow,Virtanen}:
\begin{equation}
    \label{eqn:CXELoss}
    \text{CXE}[y,y^*] = - \sum_{n=0}^{k-1} y_{n} \ln(y^*_{n})
\end{equation}
Recall a sample belongs to class $j$. Since $y$ is one-hot-encoded, the only non-zero term in the sum is $y_{j} \ln(y^*_{j})$, where $y^*_{j} \in [0,1]$.  This returns a negative number, which we multiply by $-1$ to always yield a positive cost value. When a sample produces $y^*_{j} << 1$, meaning a poor prediction, the CXE cost returns a large value and when $y^*_{j} \approx 1$, meaning a strong prediction, then the CXE cost returns a small value. We provide a visualization of this behavior in Fig. (\ref{fig:CXELoss}).

@@

ADAM provides a much more powerful and stable optimization algorithm at a higher computational cost \cite{Geron}. Note that the superscript $(i)$ indicates an iteration index where a superscript $i$ is exponentiation. Typically we initialize $\rho_1 = 0.9$, $\rho_2 = 0.999$, and introduce $\delta \approx 10^{-7}$ to avoid possible division of near-zero values \cite{Goodfellow}. All operations in the final three steps are applied element-wise. ADAM has experimentally shown to be a very effective optimizer and useful for a wide range of data sets and architectures which is why we have chosen to implement it for this task \cite{Geron,Goodfellow}.

@@

As mentioned in section \ref{subsec:structure} Eq. \ref{eqn:LayerFunction}), each layer in the neural network acts as it's own self-contained function that operates on the activations of the previous layer to transform them into the activations for the current layer. This function can take many different forms depending on where the layer sits in the chain, the type of information in the previous layer, and the nature of the solution that the neural network is trying to model \cite{Goodfellow,Loy}. For this classification  task, we employ five different layer types - (i) Dense, (ii) 2-dimensional convolution, (iii) 2-dimensional maximum-pooling, (iv) flattening, and (v) concatenation. In this section, we detail the forward-pass process of each layer and in Fig. (\ref{fig:Architecture}) we show how the combination of these layers are used to form the multimodal architecture of this network.

@@

\item\textbf{Dense Layer}
The dense layer, also called \textit{fully-connected} or \textit{affine} layer is one of earliest layer types used in artificial neural networks \cite{Geron,Loy}. A dense layer is composed of a collection of artificial neurons, each of which contains a numerical value within it (usually a floating-point number), called the \textit{activation} of that neuron. To compute the activation of any neuron, $x^{(l)_{i}}$, we take a weighted linear-combination of every neuron in the previous layer and add a constant to it, called a \textit{bias} \cite{Geron,Goodfellow,Loy}.

It then follows that a full \textit{layer} of activations, is produced from the matrix-vector product of a weight-matrix $W^{(l)}$, and the previous layers of neurons $x^{(l-1)}$, and then added to a bias $b^{(l)}$ \cite{Goodfellow,Tensorflow}. We can model this transformation as :
\begin{equation}
\label{eqn:DenseCall}
x^{(l)} = W^{(l)}x^{(l-1)} + b^{(l)}
\end{equation}
Suppose the layer $(l-1)$ contains $m$ neurons, and layer $(l)$ contains $n$ neurons. Each object then belongs to space: $x^{(l)} \in \mathbb{R}^(n \times 1)$, $x^{(l-1)} \in \mathbb{R}^(m \times 1)$, $b^{(l)} \in \mathbb{R}^(n \times 1)$, and $W^{(l)} \in \mathbb{R}^(n \times m)$. This transformation allows for an affine-transformation of activations which enables this layer to learn roughly linear decision boundaries between adjacent layers of neurons. This property of connecting every neuron in one layer to every other neuron in the previous layer is called \textit{dense connectivity} which is what gives this layer its name.

Formal implementations of this the affine transformation layer type introduce non-linearity into the hypothesis space by applying an \textit{activation} function to Eqn. (\ref{eqn:DenseCall}) \cite{Goodfellow}. This enables us to model the far more complicated decision boundaries found in real-world problems. For this classification network, we choose to use the \textit{ReLU} activation function, defined as:
\begin{equation}
\label{eqn:ReLU}
ReLU \big[ x^{(l)} \big] = \text{max} \big( 0, x^{(l)} \big)
\end{equation} 
This activation function preserves only the positive portion of each activation. \textcolor{red}{Does this Seem Out of place to you?}

@@

The Convolution layer allows for a few advantages over the dense layer. Among these are (i) \textit{spare-connectivity}: not every input neuron is connected to every single output neuron, so it can be more computationally efficient, (ii) \textit{positional invariance}: key features can be identified regardless of where the are in the input array, and (iii) \textit{Automatic feature detection}: as the weighting kernels are trained, they update as to automatically determine dominant features in the data without explicit human instruction \cite{Geron,Goodfellow,Loy}.

@@

Pooling layers are typically place after a layer or set of layers of convolution \cite{Loy}. The purpose of this is arrangement is to reduce the width of the model at the pooling layer and ensure that only the dominant features with the largest activations are preserved \cite{Geron,Loy}. This layer is considered to be "non-trainable" because it does not use any adjustable parameters in it's transformation, only a fixed procedure that cannot be updated.

\item\textbf{Flattening}
A flattening layer is used

@@

@@

The features for this branch are derived from the time-domain or frequency-domain representation of a waveform. Each feature explored represents the values computed from each audio file, which are then assembled into a 1D array-like object called a feature vector. This vector is provided as input to the MLP branch of the neural network, which can be found on right side of Fig. (\ref{fig:Architecture}). To ensure a consistency of all extracted features, all waveforms have been truncated or zero-padded to contain a consistent $M$ samples.

@@

The features for this branch are the entries in a standard 2D spectrogram representation of a waveform. A spectrogram maps the energy distribution of a signal as a function of time and frequency. It is particularly useful because it allows us to examine the frequency composition of a signal, and how that composition evolves over time. On a typical spectrogram, the passing on time is shown on the x-axis, and frequency is shown on the y-axis. Because it is represented as a 2D array or floating-point numbers, the spectrogram is effectively an image-like representation of a sound wave. Since each musical instrument produces it's own harmonic structure and transient response, each instrument can be identified by patterns within the spectrogram \cite{Virtanen}. This means than when only use this modality, the audio classification becomes very similar to an image classification task \cite{Ngiam}. We provide a few example spectrograms for reference in Fig. (\ref{fig:Example Spectrograms}).

A spectrogram is produced by applying the technique of \textit{frame-blocking} to a waveform to produce a matrix of short-time analysis frames \cite{Liu,Virtanen}. We choose each frame to contain $N = 1024$ samples, with a $768$ sample overlap between adjacent frames. 
\begin{figure}[H]
    \centering
    \includegraphics[width=0.45\textwidth]{figures/AnalysisFrames.png}
    \caption{Analysis frames in relation to a full waveform. This figures has been adapted from \cite{Liu}.}
    \label{fig:AnalysisFrames}
\end{figure}
We organize the analysis frames into a matrix, $A$ with shape $N \times k$, such that each row is a frame. We apply a Hanning window of length $N$ to each row of $A$. This eliminates discontinuities at the edges of each analysis frame, and allows for a cleaner transform into frequency-space \cite{Virtanen}. The effect of applying an Hanning window function to an analysis frame of $1024$ samples can be found in Fig. (\ref{fig:WindowFrame}). We tail-pad each time-series analysis frame by an additional $1024$ zeros which brings each frame to $2048$ samples, and gives a higher frequency-space resolution \cite{Sahidullah}. The spectrogram, $S$, is produced by computing:
\begin{equation}
    \label{eqn:Spectrogram}
    S = \big( \mathbb{W}A^T \big) \odot \big( \mathbb{W}A^T \big)^* 
\end{equation}
Where $\mathbb{W}$ is the standard \textit{Discrete Fourier Transform} (DFT) matrix, and $^*$ indicate element-wise complex conjugation. For $k$ analysis frames, matrix $S$ has shape $2048 \times k$.

\begin{figure}[h]
    \centering
    \subfigure[]{\includegraphics[width=0.45\textwidth]{figures/AnalysisFrame.png}}
    \subfigure[]{\includegraphics[width=0.45\textwidth]{figures/WindowedFrame.png}}
    \caption{(a) A standard time-series analysis frame made up of $1024$ samples and (b) The same frame with a standard Hanning Window Applied to it. This sample is taken from a bowed violin playing an $A4$ note.}
    \label{fig:WindowFrame}
\end{figure}

The standard western concert musical instruments that make up our base data set rarely have fundamental frequencies that extend beyond $6,000$ Hz. This means, that we will rarely see much energy or characteristic behavior above the $12,000$ Hz mark in the spectrograms. For this reason, we choose to crop each matrix as to include only information corresponding to the region of frequency-space form $0$ Hz to $12$ kHz. We also choose to zero pad, or crop matrix $S$ so that there are exactly $k = 256$ columns. The result is an image-like representation of a digital audio file with shape $558 \times 256$, which is passed as a feature to the input of the CNN branch of the HNN. This is the second of the two modalities that we choose to represent our data sample.

>>>>

By connecting multiple layers in succession, we form a chain-like structure of layers that describes the flow of information and allows us to model a neural network as a directed computational graph where each node is a layer that performs an operation  \cite{Geron,Goodfellow}. In a feed-forward network with $L$ layers, inputs $x^{(0)}$ and outputs $x^{(L-1)} = y^*$, we can represent a layer-chain such as:
\begin{equation}
    \label{eqn:LayerChain}
    x^{(0)} \rightarrow f^{(0)} \rightarrow f^{(1)} \rightarrow ... \rightarrow f^{(L-2)} \rightarrow f^{(L-1)} \rightarrow y^*
\end{equation}
The structure of a neural network's layer-chain is often referred to as its \textit{architecture} and chooses the \textit{hypothesis space} of the model. This is the set of all possible functions that the network can choose as a solution \cite{Goodfellow,James}. For deep neural networks this function chain can be dozens of layers long, contain upwards of millions of parameters, may enable the model to cover a much more diverse set of solutions \cite{Geron}. By carefully choosing the architecture to compliment the nature of the problem, we build a preference of solutions into the model which enables it to perform better \cite{Goodfellow}.

The hybrid neural network used in this instrument classification task differs from the more standard architecture in Eq. (\ref{eqn:LayerChain}) in that the function chain consists of two input layers, $f_a$ and $f_b$ that receive inputs $x^{(0)}_a$ and $x^{(0)}_b$ respectively. Each layer chain uses a particular arrangement of layers to form a sub-graph that compliments each type of input data. The activations are processed and transformed into compatible 1-dimensional format before they are combined at a designated hidden layer that concatenation them into a new layer. This new layer marks the start of final sub-graph, $f_c$ that leads to the output of the neural network. We represent the layer-chain structure for a generalized double-input HNN, with inputs $x^{(0)}_a$ and $x^{(0)}_b$ and outputs $y^*$ as :

@@

The combination of two sub-graphs, $f_a$ and $f_b$, into a third sub-graph, $f_c$, allows us to generate a single prediction, $y^*$ based on the behavior of both inputs. This property means that all layers in the network are part of one aggregate computational graph, one training process, and collectively form one hypothesis space. We detail the structure and function of each layer in the neural network in section (\ref{subsec:CNN}) and (\ref{subsec:MLP}).

@@

The inputs to a neural network or any machine learning algorithm are called \textit{features} or \textit{predictors} \cite{James}. These are compact, low-dimensional representations of a data sample that reflect its important characteristics \cite{Liu}. Features should be chosen as to have low variance within each class and high variance between classes. For digital audio classification of musical instruments, features of the same musical instruments should exhibit very similar properties, while features from difference musical instruments should exhibit non-similar properties. In any learning algorithm, the role of feature engineering is to translate the raw time-series information from a digital audio file into descriptors that maximize the classification performance \cite{Virtanen}. Although the neural network will classify the musical instrument within the digital audio sample, it will never interact with the waveform directly and instead will rely solely on these features. In a classification model that uses $p$ unique predictors, the features typically usually combined into a single vector-like object for each sample, and then presented as input to the neural network \cite{Geron}. We detail the classification features for this neural network in section (\ref{subsec:FeaturesMLP}) and (\ref{subsec:FeaturesCNN}).

The outputs to a classification neural network encodes the prediction that the model has made given a particular set of inputs from a sample. For a classifier with $k$ classes, there are $k$ neurons in the output layer, each one representing a particular class. These output activations which are transformed such that the sum (or $L-1$ norm) of them is identically $1$ so that they are collectively treated as a discrete probability density \cite{Goodfellow,James}. The neuron with the highest activation value indicates the class that the models predicts in input sample originated from. For classification of musical instruments, each neurons represents a musical instrument type. We use $37$ classes of instruments encapsulating woodwinds, brass, strings, mallet percussion, and synthesizers.

@@

A neural network that uses $p$ features, $k$ classes, and $b$ samples per batch is trained by providing inputs $X \in \mathbb{R}^{(b \times p)}$, with a corresponding set of labels $Y \in \mathbb{Z}^{(b \times k)}$. Consider a sample $x \in X$ that is expected to produce corresponding label $y \in Y$. When passed into the network, $x$ produces some prediction label denoted as $y^*$. For an untrained network, we anticipate that $y^*$ and $y$ may differ greatly - meaning that the prediction is very different that the expected output, which will lead to poor classification performance. Since we know each sample in the data set to contain one, and only one musical instrument, we \textit{one-hot-encode} the expected label $y$. For any sample $x^{(i)}$ that belongs to class $j$, we have a corresponding target vector $y^{(i)}$ given by:
\begin{equation}
    \label{eqn:OneHotEnc}
    y^{(i)} = \big[ y_{0} , y_{1}, ... , y_{j}, ... , y_{k-1} \big] = \big[ 0, 0, ... , 1, ... , 0 \big]
\end{equation}
The neural network will output an equivalently sized vector, $y^{*(i)}$, which all elements sum to $1$.

To quantify the difference between $y$ and $y^*$, we introduce a \textit{cost function}, $J(y,y^*)$ which is also called an \textit{objective function} or a \textit{loss function} \cite{James}. The exact cost function used can differ based on the nature of the learning model and task, but typically a multiple - category classification tasks uses a \textit{categorical cross-entropy} (CXE) function \cite{Geron,Goodfellow}. The CXE cost for a single sample is defined as \cite{Goodfellow,Virtanen}:
\begin{equation}
    \label{eqn:CXELoss}
    \text{CXE}[y,y^*] = - \sum_{n=0}^{k-1} y_{n} \ln(y^*_{n})
\end{equation}
Consider again a sample belongs to class $j$. Since $y$ is one-hot-encoded, the only non-zero term in the sum is $y_{j} \ln(y^*_{j})$. Since $y^*_{j} \in [0,1]$, then $\ln(y^*_{j}) < 0$ which we multiply by $-1$ to always yield a positive cost value. When a sample produces $y^*_{j} << 1$, meaning a poor prediction for that class, the CXE cost returns a large value and when $y^*_{j} \approx 1$, meaning a strong prediction for that category, then the CXE cost returns a small value. We provide a visualization of this behavior in Fig. (\ref{fig:CXELoss}).

@@

ADAM provides a much more powerful and stable optimization algorithm at a higher computational cost \cite{Geron}. Note that the superscript $(i)$ indicates an iteration index where a superscript $i$ is exponentiation. Typically we initialize $\rho_1 = 0.9$, $\rho_2 = 0.999$, and introduce $\delta \approx 10^{-7}$ to avoid possible division of near-zero values \cite{Goodfellow}. All operations in the final three steps are applied element-wise. ADAM has experimentally shown to be a very effective optimizer and useful for a wide range of data sets and architectures which is why we have chosen to implement it for this project \cite{Geron,Goodfellow}.

@@

As mentioned in section \ref{subsec:structure} Eq. \ref{eqn:LayerFunction}), each layer in the neural network acts as it's own self-contained function that operates on the activations of the previous layer to transform them into the activations for the current layer. The combination and arrangement of these layers define the model's hypothesis space which is the set of functions that the network can use as a solution. For this classification task we employ five different layer types detailed below. We also describe the forward-pass process of each layer and in Fig. (\ref{fig:Architecture}) and show how the combination of these layers are used to form the multimodal architecture of this network.

@@

\item\textbf{Dense Layer}
The dense layer, also called \textit{fully-connected} or \textit{affine} layer is one of earliest layer types used in artificial neural networks \cite{Geron}. It is composed of a collection of artificial neurons, each of which contains a numerical value (usually a floating-point number) called the \textit{activation} of that neuron. The activation of any neuron, is computed through a weighted linear-combination of every neuron in the previous layer with a \textit{bias} added as well \cite{Geron,Goodfellow,Tensorflow}. This property of connecting every neuron in one layer to every other neuron in the previous layer is called \textit{dense connectivity} which is what gives this layer its name.

It then follows that a full \textit{layer} of activations is produced from the matrix-vector product of a weight-matrix $W^{(l)}$, and the previous layers of neurons $x^{(l-1)}$, and then added to a bias $b^{(l)}$ \cite{Goodfellow,Tensorflow}. We can model this transformation as :
\begin{equation}
\label{eqn:DenseCall}
x^{(l)} = W^{(l)}x^{(l-1)} + b^{(l)}
\end{equation}
Suppose the layer $(l-1)$ contains $m$ neurons, and layer $(l)$ contains $n$ neurons. Each object then belongs to space: $x^{(l)} \in \mathbb{R}^(n \times 1)$, $x^{(l-1)} \in \mathbb{R}^(m \times 1)$, $b^{(l)} \in \mathbb{R}^(n \times 1)$, and $W^{(l)} \in \mathbb{R}^(n \times m)$. This transformation models an multidimensional affine-transformation which enables this layer to learn roughly linear decision boundaries between adjacent layers of neurons.

@@

The Convolution layer allows for a few advantages over the dense layer. Among these are (i) \textit{spare-connectivity}: not every input neuron is connected to every single output neuron, so it can be more computationally efficient, (ii) \textit{positional invariance}: key features can be identified regardless of where the are in the input array, and (iii) \textit{Automatic feature detection}: as the weighting kernels are trained, they update as to automatically determine dominant features in the data without explicit human instruction \cite{Geron,Goodfellow}.

@@

Pooling layers are typically places after a layer or set of layers of convolution. The purpose of this is arrangement is to reduce the width of the model at the pooling layer and ensure that only the dominant features with the largest activations are preserved \cite{Geron}. This layer is considered to be "non-trainable" because it does not use any adjustable parameters in it's transformation, only a fixed procedure that cannot be updated.

\item\textbf{Flattening}
A flattening layer is used to rearrange the shape of an array of activations without actually changing the values stored within them \cite{Tensorflow}. In particular, we use this layer to transform the 2D output of the convolution and pooling layers into 1D activations which can be passed into a dense layer. For an array of activations , $x^{(l-1)} \in \mathbb{R}^{(m \times n)}$, a flattening layer $f^{(l)}$ transforms them such that $x^{(l)} \in \mathbb{R}^{(1 \times mn)}$. This layer is also considered to be "non-trainable" because it does not use any adjustable parameters.

@@

Formal implementations of the dense layer and 2D convoltional layer types introduce non-linearity into the hypothesis space by applying an \textit{activation} function to Eq. (\ref{eqn:DenseCall}) and Eq (\ref{eqn:Conv2DCall}) \cite{Goodfellow}. For this classification network, we choose to use the \textit{ReLU} activation function, defined as:
\begin{equation}
\label{eqn:ReLU}
ReLU \big[ x^{(l)} \big] = \text{max} \big( 0, x^{(l)} \big)
\end{equation} 
Successive use of affine and convolutional transformations combine with non-linear activation functions allow a neural network to model the increasingly complicated decision-boundaries that arise in real-world problems \cite{Geron,Goodfellow}.

@@

The features for this branch are derived from the time-domain and frequency-domain representations of a waveform. Each feature explored represents the values computed from and audio file, which are then assembled into a 1D array-like object called a feature vector. This vector is provided as input to the MLP branch of the neural network, which can be found on right side of Fig. (\ref{fig:Architecture}). To ensure a consistency of all extracted features, all waveforms are assumed to have been truncated or zero-padded to contain a consistent $M$ samples.

@@

The features for this branch are the entries in a standard 2D spectrogram representation of a waveform. A spectrogram maps the energy distribution of a signal as a function of time and frequency. It is particularly useful because it allows us to examine the frequency composition of a signal, and how that composition evolves over time. On a typical spectrogram, the passing on time is shown on the x-axis, and frequency is shown on the y-axis. Because it is represented as a 2D array or floating-point numbers, the spectrogram is effectively an image-like representation of a sound wave. Since each musical instrument produces it's own spectral signature and transient response, each instrument can be identified by patterns within the spectrogram \cite{Virtanen}. This means than when only use this modality, the audio classification becomes very similar to an image classification task \cite{Ngiam}. We provide a few example spectrograms for reference in Fig. (\ref{fig:Example Spectrograms}).

A spectrogram is produced by applying the technique of \textit{frame-blocking} to a waveform to produce a matrix of \textit{short-time analysis frames} (STAF) \cite{Liu,Virtanen}. An STAF is a subsets of a waveform in the time-domain. We choose each frame to contain $N = 1024$ samples, with a $768$ sample overlap between adjacent frames. 
\begin{figure}[H]
    \centering
    \includegraphics[width=0.45\textwidth]{figures/AnalysisFrames.png}
    \caption{Analysis frames in relation to a full waveform. This figures has been adapted from \cite{Liu}.}
    \label{fig:AnalysisFrames}
\end{figure}
We organize the analysis frames into a matrix, $A$ with shape $k \times N$, such that each row is a frame. We apply a Hanning window of length $N$ to each row of $A$ which helps to eliminate discontinuities at the edges of each analysis frame and allows for a cleaner transform into frequency-space \cite{Virtanen}. The effect of applying an Hanning window function to an analysis frame of $1024$ samples can be found in Fig. (\ref{fig:WindowFrame}). We tail-pad each time-series analysis frame by an additional $1024$ zeros which brings each frame to $2048$ samples for a higher frequency-space resolution \cite{Sahidullah}. The spectrogram, $S$, is produced by computing:
\begin{equation}
    \label{eqn:Spectrogram}
    S = \big( \mathbb{W}A^T \big) \odot \big( \mathbb{W}A^T \big)^* 
\end{equation}
Where $\mathbb{W}$ is the standard \textit{Discrete Fourier Transform} (DFT) matrix, and $^*$ indicate element-wise complex conjugation. For $k$ analysis frames, matrix $S$ has shape $2048 \times k$.

\begin{figure}[h]
    \centering
    \subfigure[]{\includegraphics[width=0.45\textwidth]{figures/AnalysisFrame.png}}
    \subfigure[]{\includegraphics[width=0.45\textwidth]{figures/WindowedFrame.png}}
    \caption{(a) A standard time-series analysis frame made up of $1024$ samples and (b) The same frame with a standard Hanning Window applied to it. This sample is taken from a bowed violin playing an $A4$ note.}
    \label{fig:WindowFrame}
\end{figure}

The standard western concert musical instruments that make up our base data set rarely have fundamental frequencies that extend beyond $6,000$ Hz. Similarly, we rarely find much energy or characteristic behavior above the $12,000$ Hz mark in the spectrograms. For this reason, we choose to crop each matrix as to include only information corresponding to the region of frequency-space form $0$ Hz to $12$ kHz. Since the audio files were all sample at $44.1$ kHz, this nearly quarters the height of the spectrogram and choose to zero pad, or crop matrix $S$ so that there are exactly $k = 256$ columns. This combination of choices drastically reduces the width and complexity of the model at the entry layer saving significant computational resources. The result is an image-like representation of a digital audio file with shape $558 \times 256$, which is passed as a feature to the input of the CNN branch of the HNN. This is the second of the two modalities that we choose to represent our data sample.
